---
title: "Coexpression Gene Analysis of NR6A1"
output:
 html_document:
  author: "David McGaughey"
  date: "`r Sys.Date()`"
  theme: flatly
  toc: true
  toc_float: true
  code_folding: hide  
 html_notebook:
  author: "David McGaughey"
  date: "`r Sys.Date()`"
  theme: flatly
  toc: true
  toc_float: true
  code_folding: hide
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,  warning = FALSE,
  collapse = TRUE,
  fig.width = 12, fig.height = 8,
  comment = "#>",
  dpi=300
)
```
# tldr
NR6A1 is enriched for positive coexpression with coloboma associated genes in:
  - fetal eyeIntegration eye data (very much)
  - adult snRNA HRCA data (~0.05)
  
not enriched:
  - snRNA fetal HRCA (weakly - ~0.07)
  - adult eyeIntegration eye
  - adult eyeIntegration gtex/body
  - adult gtex body snRNA
  
# tldr2

SALL4 keeps appearing as a top correlated genes with NR6A1

# Explanation

Gene coexpression analysis is the study of correlation between two genes across numerous tissues. The idea is that genes with potentially related function will have similar patterns across many tissues / samples.

# Data sets
1. Human Retina Cell Atlas (HRCA, from Rui Chen)
  - snRNA seq
  - made into "pseudobulk" by summing up cell counts to the cell type / sample level
2. fetal HRCA (Rui Chen)
  - same processing as above
3. eyeIntegration adult eye
  - bulk
  - retina and rpe tissues, no AMD
  - only the cultured primary cells or straight tissue (e.g. no iPSC or organoids)
4. eyeIntegration fetal eye
  - bulk
  - same as above, just infant/fetal
5. eyeIntegration curated GTEx body
  - bulk
6. GTEx v8 body
  - snRNA prototype
  - psuedobulked like 1,2
  
# How to understand
Genes that are perfectly the same across samples have a correlation of 1. Perfectly different are -1. 

# Gene set enrichment testing genes

From Aman, Tiziana, Brian 2020 review. Tables 1 and 2 merged together. 



```{r, eval = FALSE}
source('src/adult_hrca_correlation.R')
source('src/fetal_hrca_correlation.R')
source('src/eiad_retina_rpe_correlation.R')
source('src/eiad_body_correlation.R')
source('src/gtex_v8_snRNA_correlation.R')
```

```{r, message = FALSE}
library(tidyverse)
adult_hrca <- read_tsv('../data/hrca_adult_nr6a1_cor.tsv.gz') %>% mutate(Gene = Gene2) %>% filter(!is.na(Gene)) %>% mutate(rank = row_number()) %>% mutate(Set = 'Adult Eye', Source = 'HRCA (snRNA)')
fetal_hrca <- read_tsv('../data/hrca_fetal_nr6a1_cor.tsv.gz') %>% mutate(Gene = Gene2) %>% filter(!is.na(Gene)) %>% mutate(rank = row_number()) %>% mutate(Set = 'Fetal Eye', Source = 'HRCA (snRNA)')
eiad_eye_adult <- read_tsv('../data/eyeIntegration_adult_tissue_eye_nr6a1_cor.tsv.gz') %>% mutate(Gene = Gene2N) %>% mutate(Set = 'Adult Eye', Source = 'eyeIntegration (bulk)')
eiad_eye_fetal <- read_tsv('../data/eyeIntegration_fetal_tissue_eye_nr6a1_cor.tsv.gz') %>% mutate(Gene = Gene2N) %>% mutate(Set = 'Fetal Eye', Source = 'eyeIntegration (bulk)')
#eiad_body <- read_tsv('../data/eyeIntegration_body_nr6a1_cor.tsv.gz') %>% mutate(Gene = Gene2N) %>% mutate(Set = 'Adult Body', Source = 'eyeIntegration (bulk)')
gtex_body <- read_tsv('../data/gtex_v8_snRNA_nr6a1_cor.tsv.gz') %>% mutate(Gene = Gene2) %>% mutate(Set = 'Adult Body', Source = 'GTEx V8 (snRNA)')
```

```{r, message = FALSE}
gtex_body_files <- list.files('../data/', full.names = TRUE)
gtex_body_files <- gtex_body_files[grepl("eyeIntegration_.*_nr6a1_cor\\.tsv\\.gz", gtex_body_files)]
gtex_body_files <- gtex_body_files[grep("adult|fetal|body", gtex_body_files,invert = TRUE)] 
eiad_gtex_tissue_cor <- purrr::map(gtex_body_files, read_tsv) %>% 
  bind_rows() %>% 
  mutate(Gene = Gene2N, Set = Tissue, Source = 'eyeIntegration/GTEx Body')
```

```{r, fig.width=10, fig.height=3}
colo <- read_csv("https://raw.githubusercontent.com/davemcg/eyeMarkers/master/lists/george_brooks_coloboma_2020.csv")

t_tester <- function(df){
  colo_set <- df %>% mutate(Coloboma = case_when(Gene %in% colo$Gene ~ 'Coloboma')) %>% 
    filter(!is.na(Coloboma))
  other_set <- df %>% mutate(Coloboma = case_when(Gene %in% colo$Gene ~ 'Coloboma')) %>% 
    filter(is.na(Coloboma))
  broom::tidy(t.test((colo_set$correlation),(other_set$correlation)))
}
```

```{r}
t_tester(adult_hrca)

adult_hrca %>% 
  mutate(Coloboma = case_when(Gene %in% colo$Gene ~ 'Coloboma'))  %>% 
  ggplot(aes(x=correlation)) + 
  geom_density() + 
  geom_point(aes(x=correlation,y=0), 
             data = . %>% filter(!is.na(Coloboma)), size = 1) +
  ggrepel::geom_label_repel(aes(x=correlation, y=0, label = Gene),
                            data = . %>% filter(!is.na(Coloboma)) %>% head(10),
                            direction = 'y', max.overlaps = Inf) +
  cowplot::theme_cowplot() + ylab('') +
  ggtitle("HRCA Coloboma Gene Coexpression with NR6A1")

adult_hrca %>% 
    mutate(Coloboma = case_when(Gene %in% colo$Gene ~ 'Coloboma Associated', TRUE ~ 'Remainder')) %>% 
  ggplot(aes(x=Coloboma,y=correlation)) + 
  geom_boxplot() + xlab("Gene Set") +
  cowplot::theme_cowplot()
```

```{r, fig.width=10, fig.height=3}

t_tester(fetal_hrca)

fetal_hrca %>% 
  mutate(Coloboma = case_when(Gene %in% colo$Gene ~ 'Coloboma'))  %>% 
  ggplot(aes(x=correlation)) + 
  geom_density() + 
  geom_point(aes(x=correlation,y=0), 
             data = . %>% filter(!is.na(Coloboma)), size = 1) +
  ggrepel::geom_label_repel(aes(x=correlation, y=0, label = Gene),
                            data = . %>% filter(!is.na(Coloboma)) %>% head(10),
                            direction = 'y', max.overlaps = Inf) +
  cowplot::theme_cowplot() + ylab('')

fetal_hrca %>% 
    mutate(Coloboma = case_when(Gene %in% colo$Gene ~ 'Coloboma Associated', TRUE ~ 'Remainder')) %>% 
  ggplot(aes(x=Coloboma,y=correlation)) + 
  geom_boxplot() + xlab("Gene Set") +
  cowplot::theme_cowplot()
```

```{r, fig.width=10, fig.height=3}
t_tester(eiad_eye_adult)
eiad_eye_adult %>% 
  mutate(Coloboma = case_when(Gene %in% colo$Gene ~ 'Coloboma'))  %>% 
  ggplot(aes(x=correlation)) + 
  geom_density() + 
  geom_point(aes(x=correlation,y=0), 
             data = . %>% filter(!is.na(Coloboma)), size = 1) +
  ggrepel::geom_label_repel(aes(x=correlation, y=0, label = Gene),
                            data = . %>% filter(!is.na(Coloboma)) %>% head(10),
                            direction = 'y', max.overlaps = Inf) +
  cowplot::theme_cowplot() + ylab('')

eiad_eye_adult %>% 
    mutate(Coloboma = case_when(Gene %in% colo$Gene ~ 'Coloboma Associated', TRUE ~ 'Remainder')) %>% 
  ggplot(aes(x=Coloboma,y=correlation)) + 
  geom_boxplot() + xlab("Gene Set") +
  cowplot::theme_cowplot()
```

```{r, fig.width=10, fig.height=3}
t_tester(eiad_eye_fetal)
eiad_eye_fetal %>% 
  mutate(Coloboma = case_when(Gene %in% colo$Gene ~ 'Coloboma'))  %>% 
  ggplot(aes(x=correlation)) + 
  geom_density() + 
  geom_point(aes(x=correlation,y=0), 
             data = . %>% filter(!is.na(Coloboma)), size = 1) +
  ggrepel::geom_label_repel(aes(x=correlation, y=0, label = Gene),
                            data = . %>% filter(!is.na(Coloboma)) %>% head(10),
                            direction = 'y', max.overlaps = Inf) +
  cowplot::theme_cowplot() + ylab('')

eiad_eye_fetal %>% 
    mutate(Coloboma = case_when(Gene %in% colo$Gene ~ 'Coloboma Associated', TRUE ~ 'Remainder')) %>% 
  ggplot(aes(x=Coloboma,y=correlation)) + 
  geom_boxplot() + xlab("Gene Set") +
  cowplot::theme_cowplot()

```




```{r}
t_tester(eiad_gtex_tissue_cor)
```
```{r}
save.image("../data/03_correlation.image.Rdata")
```